Informative Autonomous Sampling of Oceanographic Variables Using Joint Excursion Sets

### 2020-05-12 ans_to_reviewers/answers-to-reviewers.tex

\newcommand{\calcbC}{{\bf {\cal C}}}

\def\rx{{\texttt{T-REX\ }}}
\def\rxe{{\texttt{T-REX}}}
\def\du{{\texttt{DUNE}\ }}
\def\due{{\texttt{DUNE}}}

% == Maths
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage[normalem]{ulem} %to strike the words

\newcounter{reviewer}
\def\revcom{\textbf{Reviewer comment}}
\def\aecom{\textbf{Associate Editor comment}}
\def\reply{\textbf{Reply}}
\def\action{\textbf{Action}}

@@

\end{enumerate}
\par \vspace{1em}

Note that the reference to page $X$ is made relative to
the original submitted manuscript given to the reviewers.

%======================================================================
\section*{Editor}
%======================================================================
All changes requested from the editor are listed below and answered on a point-by-point basis.

%\begin{answers}

%	\item{Comment:}\label{ec1}
\vspace{5mm}
\noindent {\bf{Comment:}}

@@

Thank you for submitting your manuscript to The Annals of 
Applied Statistics.  Three referees, an Associate Editor (AE),
and I have read it carefully; reports from all four reviewers,
including particularly constructive reports from Referee 1
(R1R1.pdf; marked manuscript in R1R2.pdf) and the AE (R1AE1.pdf), 
can be downloaded via the links below. We all agree that your 
manuscript offers a potentially useful approach to estimating 
plume boundaries and a sequential design for the data collection
by autonomous underwater vehicles (AUVs). Citing from Referee 3
(R1E1.txt), "the authors have motivated the application well."
In this sense, the paper is consistent with the mission of AOAS.

However, the reports also identify serious shortcomings that 
prevent me from accepting your manuscript for publication at
this time. The reports are rather consistent in the nature of 
the problems that would need to be addressed for an AOAS publication.  
Briefly, these issues include:

%\item{1. Intended audience:}\label{r3c1}
\vspace{5mm}
\noindent {\bf{Main comment 1. Intended audience:}}

@@

As you know from reading the Manuscript Submission page of our
web site (from where you submitted your manuscript), "AOAS is 
intended to be read by all professional statisticians, as well 
as by other scientists who are interested in particular areas 
of statistical application. It is important for the abstract
and introduction to say: Why the topic is of interest, and What 
is to be shown, in as non-technical a manner as possible."
Remarks from the AE (point 1), Referee 1 (point 4), and Referee 2 
(R1R3.pdf, 1) refer to this question: "Who is the intended
audience?"  Quoting from Referee 1, "I am not clear about the 
intended audience for this paper.  Or to put it another way, 
for whom is this paper important?" The reports have some 
suggestions that may guide you in making the necessary critical 
revisions along these lines.

\reply: The comment by you, AE and reviewers made us realize that the initial submission had a motivation which was too applied. We were not selling the paper contributions in early enough, and it was not targeted sufficiently to professional statisticians. In the revised version we motivate more for statisticians. (See also comment 4 in the list of main changes above.)

\action: Our actions are to i) rewrite Introduction and Background sections to catch our intended audience and highlight our contributions, ii) restructure the presentation to highlight both statistical contributions as well as the impact such technologies and applied questions could provide for statisticians.

%\item{2. Novelty:}\label{r3c2}
\vspace{5mm}
\noindent {\bf{Main comment 2. Novelty:}}

@@

As noted in the AOAS mission statement, papers are expected
to include some degree of "methodological innovation."
In separate comments to the Editors, one reviewer noted,
"Methodological novelty is an issue with this paper.  Much of 
it is either standard Gaussian process calculations/formulae 
or sequential designs that were, in all but one case, previously 
proposed. Two things that could add a lot would be good code 
that others could run for the simulation studies like the authors' 
(if someone else were contemplating a similar experiment), and 
good code for finding the optimum design. That could lead to a 
fairly substantial impact of their work."

\reply: We realize that we did not stress the statistical novelty sufficiently in the submitted version. The main statistical novelty is in the closed form calculation for the design criteria of the integrated Bernoulli variance (IBV) associated with excursion sets (ESs) for multivariate processes. In the revised version this is extended to volumes (going beyond single location). We believe that part of the problem with the initial submission was the organization of these new results in the structure of the paper. This is hence re-structured in the revision.

\action: Our actions are to i) extend the closed form calculations to a criterion working on ES volume calculations (going beyond single location), ii) restructure the presentation to work directly on $p$-variate random fields and hence generalize the presentation, iii) restructure the closed form results to be presented in the main body, but with proofs in the new Appendix.

%\item{3. Assumptions }\label{r3c3}
\vspace{5mm}
\noindent {\bf{Main comment 3. Assumptions }}

@@

\reply: At this spatial scale there is often not so much data to infer a detailed statistical model, and this is the reason for our parsimonious construction for the real-data case. In other related situations, for instance that of larger fjord, ocean fronts, which are at a larger scale, there would usually be more data available for instance in the form of satellite data or from ensembles of computer model simulations. For such cases, we agree that it would make sense to embed more nuanced models, for instance that of larger variances at certain locations and stronger correlations in certain directions, but for our field trial here we attempted to have as few parameters as possible and had no reason to assume more complicated forms.

That said, in the updated simulation study, which we have generalized and made open-source there are opportunities and examples of more complex covariance models (non-stationary and non-separable and non-isotropic), but we do not conduct any inference for parameters in these models from ocean model data.

\action: Our actions include i) a better description of our parameter choices for the synthetic case and the specification for the real-data case, and ii) brief notes of the extended opportunities available for the associated open-access code, iii) references to related papers (including one of our own) that explored non-stationary Gaussian models for oceanographic phenomena at a larger scale than what is used here.

%\item{4. Exposition}\label{r3c4}

\vspace{5mm}
\noindent {\bf{Main comment 4. Exposition}}

@@

%\end{answers}

%======================================================================
 \section*{Associate Editor}
%======================================================================
All changes requested from the editor are listed below and answered on a point-by-point basis.

\setcounter{reviewer}{1}

%\begin{answers}

%\item{Comments to the Author}\label{ec}

%\aecom: Specific comments:

%\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

%\item[1.1.1]\label{ec1} 
\vspace{5mm}
\noindent \aecom

General: The inferential problem considered is the paper is that of estimating the “boundary” of a plume. However, the authors never define exactly what a “boundary” is. Such a definition should be provided, as well as some discussion of why determining the location of a boundary is so important. Why is it not sufficient to create a map of predictions (with an accompanying map of prediction error variances) for the variable(s) under study throughout the region of interest (a problem for which there is a huge literature on sampling design)?\par

\reply

\action: Added references \cite{adler2000excursion,adler2009random}.

%\item[1.1.2]\label{ec2} 
\vspace{5mm}
\noindent \aecom

Page 8, second paragraph of Section 4.1: The assumptions of separability and isotropy, and even stationarity, are not really necessary for the methodological development of Sections 4–6. I suggest that you wait to make these assumptions until you get to the simulation studies and example.\par

\reply We agree that they are not necessary. In the open-source that is now available, we implement non-stationary, non-separable and non-isotropic covariance models. But we do not conduct any inference for parameters. This might be possible for abundant data, but not in our small-scale resolution example.

\action: No description of modeling assumptions in the methodological sections. A better description of our parameter choices for the synthetic case and the specification for the real-data case. Brief notes of the extended opportunities available for the associated open-access code. References to related papers (including one of our own) that explored non-stationary Gaussian models for oceanographic phenomena at a larger scale than what is used here.

%\item[1.1.3]\label{ec3} 
\vspace{5mm}
\noindent \aecom

Page 9, second paragraph of Section 4.2: It has been assumed here that the measurement errors for temperature and salinity at any given location are uncorrelated. Is this a reasonable assumption? This could easily be checked by taking replicate measurements at each site.\par

\reply

\action: The assumptions are conditionally independent measurements of salinity and temperature. We believe this is realistic as they are gathered from different sensors. In the latent Gaussian random field, the variables are dependent. 
\vspace{1em}

%\item[1.1.4]\label{ec4} 
\vspace{5mm}
\noindent \aecom

General: Related to the previous point, I do not understand why sampling locations are being restricted to a regular grid. In fact, it seems unwise because it is well known (see, for example, Chapters 9 and 22 of Handbook of Environmental and Ecological Statistics, 2019, CRC Press) that the quality of inferences for spatial fields can be improved substantially by taking some
observations very close to others.\par

\reply

\action: The AUV and the sampling happens regularly at 1 Hz, the collected measurements are then assimilated to this regular grid. The sampling density is therefore higher than the grid resolution. Still, there is a limitation on restricting the sampling locations to finite number of fixed locations. However, in order to handle the computational load and resources available on the AUV this simplification is necessary. % Comment Trygve: Should we add some more here? 
\vspace{1em}

%\item[1.1.5]\label{ec5} 
\vspace{5mm}
\noindent \aecom

Page 13, last line: Specifically, what was the design? We aren’t even told what the sample size ($n_y$) was used in these calculations.\par

\reply

\action: Our actions
\vspace{1em}

%\item[1.1.6]\label{ec1} 
\vspace{5mm}
\noindent \aecom

Page 18, first paragraph of Section 7.1: The numerical values of the regression parameters given here are not meaningful without knowing the units of “West.”\par

\reply

\action: Our actions. Clarify. In a larger scale this could be improved by having auxiliary data for instance from ocean models or satellite data.

%\item[1.1.7]\label{ec7} 
\vspace{5mm}
\noindent \aecom

Page 24, first paragraph of Fig. 8: Please define “the modeled quadratic form of the residuals” more precisely. Perhaps a mathematical expression is needed for clarity.\par

\reply

\action: Our actions
\vspace{1em}

%\end{itemize}

%\item{Minor comments to the Author}\label{emc}

\vspace{5mm}
\noindent {\bf{Minor AE comments to the Authors}}
%\reply: 
\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

%\item[1.2.1]\label{emc1} 
\item Page 3, line 19: Change “bi-variate” to “bivariate”. The same change is needed on the next-
to-last line of Page 12.

\action: We have changed this in the text - page 3 and 12.

%\item[1.2.2]\label{emc2} 
\item Page 9, three lines after (10): I presume that all off-diagonal elements of R are zeros, but this
should be stated explicitly.

\action: We have changed the statement on page 9 to: ``$\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_T$ and $r^2_S$, holding the measurement error variances of temperature and salinity observations."

%\item[1.2.3]\label{emc3} 
\item Page 19, fourth line from bottom: Change ``as also" to ``as well as"

\action: We have changed this in the text - page 19.

%\item[1.2.4]\label{emc4} 
\item Page 30, second and third references listed: ``et al." is listed after the authors’ names for each
of these two references, but actually there are no additional authors beyond the two listed.

\action: we have made this change to the citations \cite{genton2015cross} and \cite{french2016credible} - page 30.
\end{itemize}

@@

%=====================================================================
\section*{Answers to referee 1}
%=====================================================================
All changes requested from Reviewer 1 are listed below and answered on a point-by-point basis.

\setcounter{reviewer}{1}

\begin{answers}

%----------------------------------------------------------------------
\item{\revcom:  4}\label{r1c4}

Adjustments and improvements towards the intended audience?

%\begin{quote}
%\frevcomment{a change in the text}
%\end{quote}

%\begin{quote}
%  \frevcomment{another change in the text} 
%\end{quote}

\item{\revcom: 4 cont.}\label{r1c4}

On including relevant code.

\item{\revcom: 5}\label{r1c5}

On shortening the paper and moving important sentences earlier. Reduced oceanographic background, while highlighting elements that reviewers did not understand about the AUV planning, waypoint graph, etc.

\item{\revcom: 6}\label{r1c6}

It is not clear why the binary classification and excursion sets are being used. Is there really a sharp boundary at the edge of a plume? Why not mapping the salinity and temperature fields instead? After all, on page 19, “achieving good predictive performance” is mentioned as a desirable feature of a good design. A lot of work has been done on that topic. And what about indicator kriging (along with logit regression) that generalizes indicator kriging that has been used to identify plumes? Modelling the p(x) is more like prohibit analysis-what is the advantage.

\item{\revcom: 7}\label{r1c7}

 The sequential sampling approaches are not novel. they are mentioned as examples in the cited paper by Bect et al. What is novel is the study of how all these methods work first in simulation studies as well as well as in the case study. That work will be a valuable contribution.

\item{\revcom:  8}\label{r1c8}

While I can see some advantages of simplicity in the IBV of Bect et al for binary data, one could have used entropy as it has information-theoretical credentials. Furthermore it could cover design both for spatial mapping as well as indicator mapping. Finally model uncertainty could be included in the overall assessment of uncertainty. The more general issue here is what objective function should be optimized in the design? Are we in a multi-attribute criterion situation?

\item{\revcom:  9}\label{r1c9}

On the issue of covariance separability, objections have always been raised when I have used it as its unrealistic. My rejoinder has always worked, that of citing the “all models are wrong”, George Box’s famous phrase. But what I was able to show in each case is a cross-validatory assessment that showed the model was “useful”. I wonder if such a model assessment could be made here to support the assumption of separability and other things as well. By the way, in spatial mapping at least, point predictors are pretty robust against mis-specification of the covariance. Where things go sideways is in the confidence bands or ellipsoids they produce. What are the error bands when modelling binary responses? Surely the predicted plumes have fuzzy boundaries.

\item{\revcom:  10}\label{r1c10}

The strategy at the bottom of page 17 seems ad hoc. Could you not use the data collected with the two ahead design to get better estimates of parameters and then with these parameters and thereafter a better myopic design?

\action: LA improved in revision! + code. %Also studied benefits of look-ahead strategies in sequential sampling in \cite{martinelli2013dynamic}.

\item{\revcom:  11}\label{r1c11}

Are these results generalizable? In other words, how much plume-to-plume variability is there? One of the problems with modelling using a lot of parameters, is there instability over replicate experiments. To put it another way which model parameters represent plume-population parameters and which represent random plume effects?

\item{\revcom:  12}\label{r1c12}

The separability assumption is to be relaxed in future work. Should publication of this case study await the proposed improvement? It would make for a much better paper.

\item{Reviewer specific comments}\label{q11}

\item[1.1.1] “joint excursion sets” in the title would be a turnoff for most statisticians. Having read the paper pretty carefully I still am not sure why they are so labelled.\par
\action: Our actions
\vspace{1em}

\item[1.1.2] Why on page 13 were those particular values chosen for the means $\mu$? Would a different choice have made a difference?\par
\action: Preliminary runs set these trends. In a larger scale, one could here use ocean model inputs or satellite data.

@@

\item[1.1.1] “joint excursion sets” in the title would be a turnoff for most statisticians. Having read the paper pretty carefully I still am not sure why they are so labelled.\par
\action: Our actions
\vspace{1em}

\item[1.1.2] Why on page 13 were those particular values chosen for the means $\mu$? Would a different choice have made a difference?\par
\action: Preliminary runs set these trends. In a larger scale, one could here use ocean model inputs or satellite data.

\vspace{1em}

\item[1.1.3] I don’t think Figure 3 conveys too much additional benefit and eliminating it for space might be considered.\par
\action: Our actions
\vspace{1em}

\item[1.1.4] Why are the ES’s called ”excursion sets? Why excursion probabilities?\par
\action: Our actions
\vspace{1em}

\item[1.1.5] I did not spot a definition for the “waypoint graph” a phrase used several times in the paper. Again points to the need to identify the audience. I doubt that most statisticians would know that phrase.\par
\action: Clarified. 
\vspace{1em}

\item[1.1.6] In statistics journals, random variables are commonly labelled by Roman characters e.g. Z, parameters in Greek e.g. $\xi$. I realize that in other scientific domains that distinction is not always made but I do recommend the change if the paper is to appear in a stats journal.\par 
\action: Our actions
\vspace{1em}

\item[1.1.7] There is not much of novelty in the Gaussian Process theory used in the paper- its pretty much standard. I suggest that the derivations stated in detail be summarized by just the results. Statisticians will know these things and users will not care.\par
\action: Our actions
\vspace{1em}

@@

\end{answers}

% \newpage

%======================================================================
\section*{Answers to referee 2}
%======================================================================
All changes requested from Reviewer 2 are listed below and answered on a point-by-point basis.

\begin{answers}
\item{\revcom:  1 - Details on AUVs}\label{r2c1}

I know very little about the AUVs you discuss, and I suspect many of your readers will be in a similar position. Additional contextual information would be helpful

@@

I know very little about the AUVs you discuss, and I suspect many of your readers will be in a similar position. Additional contextual information would be helpful

\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

\item[2.1.1] What sort of onboard computing do they have? It may not be something like an Intel 2.1 GHz, 4-thread processor, but a more informative description of something most readers relate to would be helpful. Similar information related to RAM and hard drive space would be helpful. What are the computational limits of these machines?\par
\action: Our actions
\vspace{1em}

\item[2.1.2] How long do the AUVs typically explore in one go? Is there a minimum/maximum time?\par 
\action: Max speed is limit for coverage. Time constraints for operations related to costs of man-hours in operation. Battery capacity is not a challenge for this AUV. 
\vspace{1em}

\item[2.1.3] What kind of movements can the AUVs make? Chess-like moves (horizontal, vertical, diagonal)?\par  
\action: We clarified the descriptions of a way-point graph.
\vspace{1em}

\item[2.1.4] Is there a minimum/maximum number of points you want to sample?\par
\action: Assuming a fixed time of sampling.
\vspace{1em}

\item[2.1.5] What is the size of the grid that is typically considered? i.e., what is a typical value of $n$ in Eq.(9)?\par 
\action: Only surface considered here. No 3D with depth. 
\vspace{1em}

\item[2.1.6] Does the AUV use a selection algorithm from the very beginning or is there a minimum number of initial measurements that are made before the optimization algorithm begins?\par
\action: Batch initialization. Then adaptive sampling.
\vspace{1em}

@@

\item{\revcom:  11}\label{r2c11}

p. 11, Eq. (21). The trailing $p(\bm)$ should be $\bpi(\bm)$. Should the right side of the equality be squared?

\action: Our actions

\item{\revcom:  12}\label{r2c12}

### 2020-05-13 main_new.tex
Jo: revised sect 1 and 2 (ongoing)

\begin{aug}
\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{t1,t2}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t3}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
\author{\fnms{David} \snm{Ginsbourger}\thanksref{t4}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
\and
\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5}. \ead[label=e4]{kanna.rajan@fe.up.pt}}

\affiliation[t1]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
\affiliation[t2]{Centre for Autonomous Marine Operations and Systems (AMOS).}
\affiliation[t3]{Department of Mathematical Sciences, NTNU.}
\affiliation[t4]{Uncertainty Quantification and Optimal Design group, Idiap Research Institute, Martigny, Switzerland.}
% \affiliation[t5]{Department of Engineering Cybernetics, NTNU}
\affiliation[t5]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}

@@

Monitoring the world's oceans has gained increased importance in 
light of the changing climate and increasing anthropogenic impact. A
central problem for understanding these factors is the lack of
representative data with sufficient resolution. Most of this
\emph{undersampling} can be attributed to the large spatio-temporal
variations on which ocean processes transpire, prompting the need for
effective means of sampling. By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use of autonomous platforms can be combined with statistical methods to pursue measurements with high scientific relevance. The potential that lies in using information-based methods \citep{das11b,Das2015,fossum18b,fossuminformation}, which we highlight in this paper, is to do efficient environmental sampling and overcome three common challenges, namely:

@@

\begin{itemize}

\item Full numerical ocean models cannot provide accurate results
  online if run on robotic sensing platforms, as onboard computers
  cannot deliver the computational power required. Hence statistical
  proxy models of the environment must be used for learning where to sample.

\item With limited available information about the state of the ocean, there is substantial value in reacting to
  information obtained from measurements taken in-situ. This
  acquired information must be assimilated into statistical models that
  can be used to inform decisions on where to sample
  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{ause2008phd}.

\item For sampling problems related to environmental sensing, the
  number of choices (i.e. locations, trajectories, and candidate
  designs) is enormous, creating a
  trade-off between optimization (finding the most resource-efficient
  design to collect necessary data) and computability (arriving at a
  solution in reasonable time). To successfully resolve features, this
  trade-off has to be considered in development and practice.

\end{itemize}

Addressing this, the combination of statistical tools and robotic platforms is a
natural symbiosis which enables information-based sensing. Central to
this is the ability to model spatially-correlated variables and
provide formal measures of uncertainty. Our formulation is based on
Gaussian Processes (GPs) as they allow efficient implementation 
and evaluation in real time onboard a robotic platform.

@@

In this work, we build on joint excursion sets (ES) and excursion
probabilities (EP) for the spatial design of experiments. The goal is
to design sampling strategies for reducing the spatially-integrated
Bernoulli variance (IBV) in the ES indicators. Our main contributions
include: i) extending the IBV criteria to multivariate
GPs, thereby allowing the evaluation to include multiple environmental
parameters such as temperature and salinity; ii) developing methods
for information-based environmental sensing and autonomous data
collection; and iii) demonstrating the approach in the real-world
scenario of sampling a river plume using an AUV.

@@

Central to understanding the changes taking place in the upper
water-column is an understanding of biophysical interaction driven by
an agglomeration of physical forcing (wind, topography, bathymetry,
tidal influences, and local geography, such as fresh water inflows) and
incipient micro-biology driven by planktonic and coastal anthropogenic
input, such as pollution and agricultural runoff transported into the ocean by rivers and streams. These often result in a range of ecosystem-related phenomena such as blooms, anoxic zones and plumes, all of which have significant direct and indirect effects on society.

% On observation and why robotics
Traditional data collection at sea has typically been based on static
buoys, Lagrangian floats, or ship-based methods, with significant
logistical limitations that directly impact coverage and sampling
resolution. Modern methods using satellite remote-sensing provide
large-scale coverage but have limited resolution, are limited to
sensing the surface of the ocean, and are impacted by cloud cover. The
advent of mobile robotic platforms \citep{Bellingham07} has
resulted in significant contributions to environmental monitoring and
sampling. In particular, autonomous underwater vehicles (AUVs), have
advanced the state of sampling and consequently have made robotics an
integral part of ocean observation. The previously listed challenges related to undersampling have also been driving ocean sensing practices towards more autonomous, networked, and distributed use of robotic assets (See Fig. \ref{fig:envir1}), where observation is guided by assimilation of recent measurements and statistical models.

@@

\begin{figure}[!h] 
  \centering 
  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
    0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
  \hfill
  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
  \caption{\ref{fig:envir1} Traditional ocean observation based on
    ship-based sampling has recently been augmented by autonomous and
    collaborative networked operations of robotic vehicles to resolve
    numerous bio-geochemical processes and their interactions. AUV platforms are an integral part of this network being able to reason and make decisions for efficient onboard adaptive sampling using the sense-plan-act control approach to autonomous control.
    \ref{fig:nidelven} The interaction of river and ocean creates
    processes that are challenging to map, where the combination of
    statistics and robotics can play a vital role in enabling more
    effective observation practices.}
\label{fig:envir}
\end{figure}

The application that is of focus in this work is towards spatial characterization of a frontal system specifically, fronts generated by river
plumes. Fig. \ref{fig:nidelven} shows the survey area (the outlet of
the Nidelva river, Trondheim, Norway) where cold freshwater enters
from the river, creating a strong gradient in both temperature and
salinity. Because of the local topography and the Coreolis force
\citep{coriolis1835memoire} the cold fresh water tends to flow near
land to the east. Depending on the river discharge, tidal effects,
wind, and temperature differences, this boundary often gets
distorted. Initial knowledge about the location and evolution of these
features is therefore highly uncertain, making deterministic planning
challenging. River plumes belong to a class of ocean processes that are local and
act on smaller \emph{sub-mesoscale} extent (from $5 m^2$ -- $10 km^2$) where
lateral spatial variability tends to dominate. Remote sensing and
synthetic ocean models often find it challenging to provide detail at
this scale \citep{Lermusiaux:2006} and the principal way to resolve
for fine resolution is by direct in-situ observation. At larger
\emph{mesoscale} ($>50 km^2$), both three-dimensional space and time
dynamics are important and can shift substantially, while in the case
of river plumes one can often limit the scope to cover only lateral
spatial elements. However, the methodological framework can be
extended to higher dimensional situations in a similar manner.

@@

>>>>

\begin{aug}
\author{\fnms{Trygve Olav} \snm{Fossum}\thanksref{t1,t2}, \corref{} \ead[label=e1]{trygve.o.fossum@ntnu.no}}
\author{\fnms{Cédric} \snm{Travelletti}\thanksref{t3}, \corref{} \ead[label=e1]{cedric.travelletti@stat.unibe.ch}}
\author{\fnms{Jo} \snm{Eidsvik}\thanksref{t4}, \ead[label=e2]{jo.eidsvik@math.ntnu.no}}
\author{\fnms{David} \snm{Ginsbourger}\thanksref{t3}, \ead[label=e3]{david.ginsbourger@idiap.ch}}
\and
\author{\fnms{Kanna} \snm{Rajan}\thanksref{t5}. \ead[label=e4]{kanna.rajan@fe.up.pt}}

\affiliation[t1]{Department of Marine Technology, The Norwegian University of Science and Technology (NTNU), Trondheim, Norway.} 
\affiliation[t2]{Centre for Autonomous Marine Operations and Systems (AMOS).}
\affiliation[t3]{Institute of Mathematical Statistics and Actuarial Science, University of Bern, Switzerland.}
\affiliation[t4]{Department of Mathematical Sciences, NTNU.}
\affiliation[t5]{Underwater Systems and Technology Laboratory, Faculty of Engineering, University of Porto, Portugal.}

@@

Monitoring the world's oceans has gained increased importance in 
light of the changing climate and increasing anthropogenic impact. A
central problem for understanding these factors is the lack of
representative data with sufficient resolution. Most of this
\emph{undersampling} can be attributed to the large spatio-temporal
variations on which ocean processes transpire, prompting the need for
effective means of sampling. By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use of new technologies including autonomous platforms can be combined with statistical methods to pursue measurements with high scientific relevance. In this paper we build upon the potential that lies in using autonomous underwater vehicles (AUVs) to conduct efficient sampling in the ocean \citep{das11b,Das2015,fossuminformation,fossum18b}. Addressing this, the combination of statistical tools and robotic platforms is a natural symbiosis which enables information-based environmental sensing.

There are several obstacles when doing sampling with an AUV in the ocean, and we discuss some of these to motivate our setting, and also the modeling and methodological choices that will be made. First, full numerical ocean models based on complex physical processes cannot provide accurate results for online operations because the onboard computers cannot deliver the computational power required. Hence statistical proxy models of the environment must be used on the AUV. For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest. Second, with limited available information about the state of the ocean, one must usually plan for a targeted operation where a relevant criterion is used to extract the most valuable designs, preferable in such a way that there is substantial value in reacting to information obtained from measurements taken in-situ. The acquired information must then be assimilated into the statistical model onboard the AUV, so that it can be used to re-compute the sampling criterion and in this way actively inform decisions on where to sample sequentially. We use a criterion focusing on the prediction of excursion sets (ESs) for multivariate processes of interest. Third, the number of sampling choices (i.e. locations, trajectories, and candidate
designs) is enormous, creating a trade-off between optimization (finding the most resource-efficient design to collect necessary data) and computability (arriving at a
solution in reasonable time). One must often compromise in both development and practical operations, and we will rely on short-horizon optimization routines among sequential designs that encapsulate nearest neighbors sampling in a waypoint graph created for the AUV operation.

@@

%\begin{itemize}
%
%\item Full numerical ocean models cannot provide accurate results
%  online if run on robotic sensing platforms, as onboard computers
%  cannot deliver the computational power required. Hence statistical
%  proxy models of the environment must be used for learning where to sample.
%
%\item With limited available information about the state of the ocean, there is substantial value in reacting to
%  information obtained from measurements taken in-situ. This
%  acquired information must be assimilated into statistical models that
%  can be used to inform decisions on where to sample
%  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{ause2008phd}.
%
%\item For sampling problems related to environmental sensing, the
%  number of choices (i.e. locations, trajectories, and candidate
%  designs) is enormous, creating a
%  trade-off between optimization (finding the most resource-efficient
%  design to collect necessary data) and computability (arriving at a
%  solution in reasonable time). To successfully resolve features, this
%  trade-off has to be considered in development and practice.
%
%\end{itemize}

%Addressing this, the combination of statistical tools and robotic platforms is a
%natural symbiosis which enables information-based sensing. Central to
%this is the ability to model spatially-correlated variables and
%provide formal measures of uncertainty. Our formulation is based on
%Gaussian Processes (GPs) as they allow efficient implementation 
%and evaluation in real time onboard a robotic platform.

@@

We build on ESs and excursion
probabilities (EPs) for the spatial design of experiments. The goal is
to design sampling strategies for reducing the spatially-integrated
Bernoulli variance (IBV) in the ES indicator variables. 
Recent work on ESs and EPs connected to spatial statistics include
\cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}.
Our focus is on uncertainty
characterization and how it is influenced by sampling. In this sense,
our work is similar to \cite{bect2012}, \cite{chevalier2014fast}, and
\cite{azzimonti2016quantifying} who describe analytical results for
the IBV for univariate processes. A major contribution of our work is
to derive closed-form results for this design criteria for situations
where the underlying model is based on multivariate GPs. Another is that we embed this approach in a sequential strategy that runs onboard an AUV, and in applying this in the real-world
scenario of autonomously sampling temperature and salinity variables in a river plume.

Motivating examples relevant for ESs of multivariate processes are abundant. In medicine, doctors do not solely rely on a single symptom but must see several combined effects before making a diagnosis. In our context of spatial and environmental applications, a famous example is that of photosynthesis where plants and algae produce oxygen only when given enough of every component; light, carbon dioxide and water. Our key motivation in this paper is that temperature and salinity, combined, drive several processes in the oceans. For instance, salmon lice larvae are known to depend on relatively high temperature and high salinities to survive. Salmon lice are parasites to salmon, and the bivariate ES for temperature and salinity is hence important to understand for sustainable activities in salmon fisheries. The salinity and temperature excursions of a river plume similarly characterize its oceanographic and bio-geochemical processes \cite{hopkins2013detection,Pinto2018}, and sampling should be aimed at mapping relevant parts to reduce uncertainty in the temperature and salinity ES.

@@

Central to understanding the changes taking place in the upper
water-column is an understanding of biophysical interaction driven by
an agglomeration of physical forcing %(wind, topography, bathymetry,
%tidal influences, and local geography, such as fresh water inflows) 
and incipient micro-biology driven by planktonic and coastal anthropogenic
input, such as pollution and agricultural runoff transported into the ocean by rivers and streams. These often result in a range of ecosystem-related phenomena such as blooms%, anoxic zones
and plumes. %, all of which have significant direct and indirect effects on society.

% On observation and why robotics
Traditional data collection at sea has typically been based on static
buoys, floats, or ship-based methods, with significant
logistical limitations that directly impact coverage and sampling
resolution. Modern methods using satellite remote-sensing provide
large-scale coverage but have limited resolution, are limited to
sensing the surface of the ocean, and are impacted by cloud cover. The
advent of mobile robotic platforms \citep{Bellingham07} has
resulted in significant contributions to environmental monitoring and
sampling. In particular, AUVs have
advanced the state of sampling and consequently have made robotics an
integral part of ocean observation. The previously listed challenges related to undersampling have also been driving ocean sensing practices towards more autonomous, networked, and distributed use of robotic assets (See Fig. \ref{fig:envir1}), where observation is guided by assimilation of recent measurements and statistical models.

@@

\begin{figure}[!h] 
  \centering 
  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
    0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
  \hfill
  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
  \caption{\ref{fig:envir1} Traditional ocean observation based on
    ship-based sampling has recently been augmented by autonomous and
    collaborative networked operations of robotic vehicles to resolve
    numerous bio-geochemical processes. % and their interactions. 
    AUV platforms are an integral part of this network being able to reason and make decisions for efficient onboard adaptive sampling. % using the sense-plan-act control approach to autonomous control.
    \ref{fig:nidelven} The interaction of river and ocean creates
    processes that are challenging to map, where the combination of
    statistics and robotics can play a vital role in enabling more
    effective observation practices.}
\label{fig:envir}
\end{figure}

The application that is of focus in this work is towards spatial characterization of a frontal system specifically, fronts generated by river
plumes. Fig. \ref{fig:nidelven} shows the survey area (the outlet of
the Nidelva river, Trondheim, Norway) where cold freshwater enters
from the river, creating a strong gradient in both temperature and
salinity. Because of the local topography and the Coreolis force
%\citep{coriolis1835memoire} 
the cold fresh water tends to flow near
land to the east, %. Depending on the river discharge, tidal effects,
%wind, and temperature differences, this boundary often gets
%distorted. Initial knowledge about the location and evolution of these
but the locations of these features are highly uncertain, making deterministic planning
challenging. 
%River plumes belong to a class of ocean processes that are local and
%act on smaller \emph{sub-mesoscale} extent (from $5 m^2$ -- $10 km^2$) where
%lateral spatial variability tends to dominate. 
Remote sensing and
synthetic ocean models often find it challenging to provide detail at
this scale \citep{Lermusiaux:2006} and the principal way to resolve
for fine resolution is by direct in-situ observation. 
%At larger \emph{mesoscale} ($>50 km^2$), both three-dimensional space and time
%dynamics are important and can shift substantially, while in the case
At this scale of river plumes one can further often limit scope to covering only lateral near surface
spatial elements, and the case appears very amenable for AUVs usage. 
%However, the methodological framework can be extended to higher dimensional situations in a similar manner.

@@

\cite{adler2000excursion,adler2009random}.
